---
title: "Manipulating Spatial Data"
output: statsr:::statswithr_lab
---
<div id="Introduction">
This tutorial aims to explain some **useful functions for manipulating spatial data using some packages**.

</div>

#### WORKING WITH SPATIAL DATA

First, let's start by loading the required packages:
```{r message=FALSE, warning=FALSE}
library(rgdal)
library(data.table)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(htmlwidgets)
```

#### ACCESSING THE DATA
  One way to read spatial data is to use `readOGR()` function from `rgdal` package. the `(.)` refers to the `current working directory` where the shapefile is stored. The second argument refers to `the name of the shapefile whitout the extension.
  
```{r}
metro <- readOGR(".", "metro_station")
```

```{r}
summary(metro)
```

  The summary shows that the shapefile is a `SpatialPointsDataFrame`, followed by the `Coordinate System` information, `proj4string` and finally the `data frame`.

  Another way to read shapefiles is to use `readShapeSpatial("metro_station")` function from `maptools` package. This latter doesn't display the `Coordiante System (CS)` information.

  The function `read_sf("metro_station")` from the `sp` package allows to read shapefiles and displays the data as a tibble.  

  The `@` allows to access the data frame like bellow. Another way to access the data frame is `data.frame(metro)`. 
The name of some columns has changed in order to be reconized by the package leaflet later.
The data frame below shows the 4 most important variables or columns:

  + `nom` refers to the name of metro stops;
  + `ligne` refers to the metro line;
  + `lon` refers to the longitude of each stop;
  + `lat` refers to the latitude of each stop.

```{r echo=FALSE}
metro_df <- metro@data
setnames(metro_df, old = c("x_wgs84", "y_wgs84"), new = c("lon", "lat"))
head(metro_df)
```

#### MAPPING SPATIAL DATA
  Many packages have been developped to map spatial data. `leaflet` is an open source JavaScript library for creating interactive maps. 
leaflet uses tiles. Tiled web maps join many map images together and when a user zooms or pans an interactive leaflet map, new tiles are fetched as needed to provide the requested view of the map.

In the code below, an html widget has been initialized with `leaflet()` function call. Notic that leaflet leverages the pipe operator that is common in the `tidyverse`. This allows to chain a series of functions together. 
`addTiles()` adds a map tile and by adding `addMarkers()`, the web map will place a blue drop pin at the coordinates with the specific name(nom) shown when pointing into a particular coordinate.
leaflet() reconizes automatically the lon at lat in the data frame.

```{r fig.width=10, fig.height=8, message=FALSE, warning=FALSE}
metro_df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = ~nom)
```

As shown above, all the points are displayed and there is no way to select just one type of metro line.
To be more flexible, we subset the data based on a particular metro line by using `filter(ligne == "x")` and add each in the `addCircleMarkers()` as shown later. In this case, the full data frame will be removed from the code.

```{r echo=FALSE}
ligne_a <- metro_df %>%
  filter(ligne == "a")
```

```{r echo=FALSE}
ligne_b <- metro_df %>%
  filter(ligne == "b")
```

The leaflet package comes with `more than 100 map tiles`. These tiles are stored in a list called `providers` and can be added to a map using `addProviderTiles() instead of addTiles()`.
`addCircleMarkers()` id very useful when we have many points to display and can replace `addMarkers()` function.
`addLayersControl()` adds a control map to select the metro line we like to visualize or a particular tile. 
```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=8}
pal <- colorFactor(palette = c("darkgreen", "darkblue"), levels = c("a", "b"))

m <- leaflet() %>%
  addSearchOSM() %>% 
  addTiles(group = "OSM") %>%
  
  
  addProviderTiles("CartoDB", group = "Carto") %>%
  addProviderTiles("Esri", group = "Esri") %>%
  addCircleMarkers(data = ligne_a, radius = 5, col = ~pal(ligne), label = ~htmlEscape(nom), group = "a") %>%
  addCircleMarkers(data = ligne_b, radius = 5, col = ~pal(ligne), label = ~htmlEscape(nom), group = "b") %>% 
  addLayersControl(baseGroups = c("OSM", "CartoDB", "Esri"), overlayGroups = c("a", "b")) %>%
  addLegend(pal = pal, values = c("a", "b"), opacity = 0.5, title = "Ligne", position = "topright") 

m
```

After storing a leaflet map in an object, the `saveWidget()` function from `htmlwidgets` allows to write the map to an html file and store it locally or on a server.
```{r}
saveWidget(m, file = "metroMap.html")
```






